### 2020-03-09 latex/Image_b_convex.tex
Multiple small corrections
overview.tex:
spelling mistakes and minor notation changes
Changed the second example to make it more reasonable
TODOs: Abstract, Introduction, Formulation LPs, Numerical results, Conclussion

Bib_RK_optim.bib:
Hairer as complete book 

Image_b_direct.tex:
Notation change

Image_b_convex.tex:
Notation change, 
Make it fit into context of direct adaptation

\begin{tikzpicture}
    % Draw axes

	\coordinate (bo) at (1.2,1.8);
    \coordinate (b) at (2,1);
    \coordinate (db) at ($(b)-(bo)$);
    \coordinate (b1) at ($0.4*(db)+(b)$);
    \coordinate (b0) at ($-0.2*(db)+(bo)$);;

    \draw [<->,thick] (0,2.5) node (yaxis) [above] {$b_2$}
        |- (3,0) node (xaxis) [right] {$b_1$};

%\draw[thick,red,*-*] (b0) -- (b1);
   \node[circle,inner sep=1.5pt,draw,fill,red] (A) at (b0) {};
   \node[circle,inner sep=1.5pt,draw,fill,red] (B) at (b1) {};
   \draw[thick,red,-] (b0) -- (b1);

	\draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$b_{a}$} (b0);
    \draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$\tilde{b}$} (b);
    \draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8] {$b_{b}$} (b1);

\end{tikzpicture}