### 2020-03-31 nusslein_vsrp/latex/Image_b_direct.tex

\begin{tikzpicture}
    % Draw axes
	\coordinate (borig) at (1.2,1.8);
    \coordinate (b) at (2,1);
	\coordinate (db) at ($(b)-(borig)$);

    \draw [<->,thick] (0,2.5) node (yaxis) [above] {$b_2$}
        |- (3,0) node (xaxis) [right] {$b_1$};

%Add subspace 
	\draw[thick,red,dotted] ($-0.9*(db)+(borig)$) -- ($0.9*(db)+(b)$);
	\draw[thick,red] ($-0.6*(db)+(borig)$) -- ($0.6*(db)+(b)$);

	\draw[thick,->] (borig) -- node[anchor=south west ,pos = 0.5] {$\Delta b$} (b);
	\draw[thick,->] (0,0) -- node[anchor=south east,pos = 0.8] {$b$} (borig);
    \draw[thick,->] (0,0) -- node[anchor=north west,pos = 0.8] {$\tilde{b}$} (b);

\end{tikzpicture}